---
title: "HNSCC HPV Cohort Subsets"
author: "Aurora S Blucher"
date: "7/26/2019"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(tidyr)
library(stringr)
```

## HNSCC HPV SUBSETS

This script uses the HPV consensus annotation file to generate HPV-positive and HPV-negative subsets from the full TCGA Head and Neck Squamous Cell Carcinoma cohort. As part of the code review, we check to make sure that the HPV subsets we generate match previous HPV subsets. 

These HPV cohort subsets are then used in another notebook in this project to find HPV-status cohort specific light and dark pathways.

```{r}
setwd("/Users/blucher/Box/OHSU (blucher@ohsu.edu)/HNSCC/gabby_manuscript/submission_PLOS/Figures_AndInput_FromGabby")

################
###DATA
################

#hnscc mutation data, same file used for pathway analysis of full cohort
hnscc_mutation<-read.csv(file="TCGA.HNSC.mutect.84c7a87a-9dcc-48fb-bd69-ba9d6e6f3ca2.DR-7.0.somatic_cleaned.maf",header = TRUE, sep="\t")
hnscc_mutation_cleanedBarcode<-hnscc_mutation %>% mutate(Patient_barcode=str_sub(Tumor_Sample_Barcode, 0, 12))
hnscc_mutation_patientIDs<-hnscc_mutation_cleanedBarcode %>% select(Tumor_Sample_Barcode, Patient_barcode) 
head(hnscc_mutation_patientIDs)

#consensus hpv annotation file from Mitzi Boardman
hpv_annotation_mitzi<-read.csv(file="HPV_Annotation_MB.csv", header = TRUE, sep=",")
head(hpv_annotation_mitzi)
```

## HPV POSITIVES

The HPV Positive cohort is created by subsetting our fill TCGA MAF file to include only those TCGA samples with an HPV annotation of "Highest Confidence Positive".

```{r}
#positives= final call is either highest confidence positive or high confidence positive
#using highest confidence positive only
hpv_positivesM<-hpv_annotation_mitzi %>% filter (FINAL_HPV_CALL=="Highest Confidence Positive" )
n_distinct(hpv_positivesM) #62 highest confidence positives

#how many positives are in the tcga hnscc file from gabby? 
#ASB note, there may be more patient IDs in annotation file due to using a later version than when Gabby pulled MAF
hpv_posM_inourset<-hnscc_mutation_patientIDs %>% filter(Patient_barcode %in% hpv_positivesM$TCGA_BARCODE) %>% unique()
n_distinct(hpv_posM_inourset)#57 HPV positive samples in our tcga hnscc file


#CREATE HPV POSITIVE SUBSET MAF 
#create HPV positive subset maf, then remove our cleaned barcode
hpv_pos_subset_for_maf<-hnscc_mutation_cleanedBarcode %>% filter(Patient_barcode %in% hpv_positivesM$TCGA_BARCODE) %>% select(-Patient_barcode, "patientid"=TCGA_Patient_Barcode) %>%unique()
View(hpv_pos_subset_for_maf)
levels(factor(hpv_pos_subset_for_maf$Variant_Classification)) #9 levels here
write.table(hpv_pos_subset_for_maf, "TCGA.HNSC.mutect.84c7a87a-9dcc-48fb-bd69-ba9d6e6f3ca2.DR-7.0.somatic_cleaned_HPVpos_checked.maf", sep="\t", row.names = FALSE, col.names = TRUE, quote = FALSE)


#COMPARING TO HPV POS SUBSET MAF FILE FROM GABBY
#note, our maf file is larger (8013 rows compared to 5194 rows, same 121 cols)
#check to make sure these are the same 57 samples as in Gabby's split files

#HPV positive cohort from Gabby
hpvPos_fromSubsetFile<-read.csv(file="TCGA.HNSC.mutect.84c7a87a-9dcc-48fb-bd69-ba9d6e6f3ca2.DR-7.0.somatic_cleaned_HPVpos.maf",header = TRUE, sep="\t")
View(hpvPos_fromSubsetFile)
hpvPos_fromSubsetFile_checkunique<-hpvPos_fromSubsetFile %>% unique()
n_distinct(hpvPos_fromSubsetFile$Tumor_Sample_Barcode) #57 positive samples used in the analysis
hpv_posM_inourset_check<-hpv_posM_inourset %>% mutate("Annotated_fromMitzi"=ifelse(hpv_posM_inourset$Tumor_Sample_Barcode %in% hpvPos_fromSubsetFile$Tumor_Sample_Barcode, "Yes", "No"))
#View(hpv_posM_inourset_check) #yes, these are the same 57 samples


#it seems to be the variant classification designation
#17 levels of variant classification in Gabby's file
levels(factor(hpvPos_fromSubsetFile$Variant_Classification)) #17 

#9 levels of variant classification in our file
variant_Classification_List<-levels(factor(hpv_pos_subset_for_maf$Variant_Classification))

#filter gabby's file to our 9 variant classifications -> produces the same 5194 rows
filtering_try<-hpvPos_fromSubsetFile %>% filter(Variant_Classification %in% variant_Classification_List)
#View(filtering_try)
#View(hpv_pos_subset_for_maf)

#CHECK DATA FRAMES ARE EQUAL
#try all_equal from dplyr
#incompatible types distance/ somatic - just remove these columns for now, they are not used
filtering_try_tocompare<-filtering_try %>% select(-DISTANCE, -SOMATIC)
hpv_pos_subset_for_maf_tocompare<-hpv_pos_subset_for_maf %>% select(-DISTANCE, -SOMATIC)
#ignore column order, ignore row order
check<-all_equal(filtering_try_tocompare, hpv_pos_subset_for_maf_tocompare, ignore_col_order = TRUE, ignore_row_order = TRUE, convert=TRUE)
check #YES DATA FRAMES ARE THE SAME
```

## HPV NEGATIVES

The HPV Negative cohort is created by subsetting our fill TCGA MAF file to include only those TCGA samples with an HPV annotation of "High Confidence Negative".

```{r}

#positives from mitzi - final call is either highest confidence positive or high confidence positive
#using highest confidence positive only
hpv_negativesM<-hpv_annotation_mitzi %>% filter (FINAL_HPV_CALL=="High Confidence Negative" )
#View(hpv_negativesM)#121 high confidence negatives here

#how many negatives are in the tcga file from gabby?
hpv_negM_inourset<-hnscc_mutation_patientIDs %>% filter(Patient_barcode %in% hpv_negativesM$TCGA_BARCODE) %>% unique()
#View(hpv_negM_inourset) #118 HPV negative samples in Gabby's set

#CREATE HPV NEGATIVE SUBSET MAF 
#create HPV NEGATIVE subset maf, then remove our cleaned barcode
hpv_neg_subset_for_maf<-hnscc_mutation_cleanedBarcode %>% filter(Patient_barcode %in% hpv_negativesM$TCGA_BARCODE) %>%select(-Patient_barcode, "patientid"=TCGA_Patient_Barcode) %>%unique()
nrow(hpv_neg_subset_for_maf) #12771 rows
write.table(hpv_neg_subset_for_maf, "TCGA.HNSC.mutect.84c7a87a-9dcc-48fb-bd69-ba9d6e6f3ca2.DR-7.0.somatic_cleaned_HPVneg_checked.maf", sep="\t", row.names = FALSE, col.names = TRUE, quote = FALSE)


#COMPARING TO HPV NEG SUBSET MAF FILE FROM GABBY
#note, our maf file is larger (8013 rows compared to 5194 rows, same 121 cols)
#check to make sure these are the same 57 samples as in Gabby's split files

#check to make sure these are the same 118 samples as in Gabby's split files
#HPV negative cohort
hpvneg_fromSubsetFile<-read.csv(file="TCGA.HNSC.mutect.84c7a87a-9dcc-48fb-bd69-ba9d6e6f3ca2.DR-7.0.somatic_cleaned_HPVneg.maf",header = TRUE, sep="\t")
n_distinct(hpvneg_fromSubsetFile$Tumor_Sample_Barcode) #118 positive samples used in the analysis
hpv_negM_inourset_check<-hpv_negM_inourset %>% mutate("Annotated_fromMitzi"=ifelse(hpv_negM_inourset$Tumor_Sample_Barcode %in% hpvneg_fromSubsetFile$Tumor_Sample_Barcode, "Yes", "No"))
#View(hpv_negM_inourset_check) #yes- 118 samples
nrow(hpvneg_fromSubsetFile) #19058 rows

#it seems to be the variant classification designation
levels(factor(hpvneg_fromSubsetFile$Variant_Classification)) #18 here

#9 levels of variant classification in our file
levels(factor(hpv_neg_subset_for_maf$Variant_Classification))
variant_Classification_List_neg<-levels(factor(hpv_neg_subset_for_maf$Variant_Classification))

#filter gabby's file to our 9 variant classifications -> produces the same 5194 rows
filtering_try_neg<-hpvneg_fromSubsetFile %>% filter(Variant_Classification %in% variant_Classification_List_neg)
nrow(filtering_try_neg) #12774
#View(filtering_try)
nrow(hpv_neg_subset_for_maf)#12771

#NOTE, NOT EXACTLY EQUAL!
#WE HAVE 3 ADDITIONAL ROWS IN THE SUBSET FILE FROM GABBY THAT ARE NOT IN OUR HPV NEG SUBSET FILE

#CHECK DATA FRAMES ARE EQUAL
#try all_equal from dplyr
#incompatible types distance/ somatic - just remove these columns for now, they are not used
filtering_try_tocompare_neg<-filtering_try_neg %>% select(-DISTANCE, -SOMATIC)
hpv_neg_subset_for_maf_tocompare<-hpv_neg_subset_for_maf %>% select(-DISTANCE, -SOMATIC)
#ignore column order, ignore row order
check_neg<-all_equal(filtering_try_tocompare_neg, hpv_neg_subset_for_maf_tocompare, ignore_col_order = TRUE, ignore_row_order = TRUE, convert=TRUE)
check_neg #DIFFERENT NUMBER OF ROWS


#use anti_join to get those 3 rows
extraRows_neg<-anti_join(filtering_try_tocompare_neg, hpv_neg_subset_for_maf_tocompare)
nrow(extraRows_neg) #3 extra rows
#notes
#all 3 data points are Variant type DELETE, but those do not appear to be excluded
#

#save this file to show that there is a difference
write.table(extraRows_neg, "TCGA.HNSC.mutect.84c7a87a-9dcc-48fb-bd69-ba9d6e6f3ca2.DR-7.0.somatic_cleaned_HPVneg_checked_3DifferingRows_072919.maf", sep="\t", row.names = FALSE, col.names = TRUE, quote = FALSE)



```
