---
title: "Figure4- Visualization of Nephrin Interactions and Drug Target Evidence"
author: "Aurora S Blucher"
date: "7/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)


```


This script creates the table for drug-target interactions for drugs interacting with the Nephrin Interaction pathway. It parses the Reactome gene set file, selects the member genes for the Nephrin Interactions pathway, and selects all drug-target interactions from the Cancer Targetome.

Produces the table "NephrinInteractions_GeneMembers_withDrugEvidence.txt" which was cross-checked with drug information pulled from ReactomeFIViz Cytoscape app. Target information in this file was used to manually adjust colors of nodes and edges in visualization of pathway for Figure 4.


```{r cars}
#PARSING REACTOME PATHWAY FILE
#Pathway->Gene members
#Using same version as Gabby's analyses (2015 version)
#read in as one column, then split
reactomePathway_GeneList<-read.csv(file="ReactomePathways 2015.02.17 13.46.25.txt",header = FALSE,sep = "\n")

#first separate pathway name and "Reactome Pathways"
#then separate_rows to convert to LONG format
#remove unneeded "ReactomePathways" indicator
pathwayList_GenesOneColumn<-reactomePathway_GeneList%>% 
  separate(V1, c("Pathway", "Extra_AndGenes"), sep = "\t", extra = "merge") %>%
  separate_rows(Extra_AndGenes) %>% filter(Extra_AndGenes!="ReactomePathways") %>% select(Pathway, "Gene"=Extra_AndGenes)

#FILTER TO NEPHRIN INTERACTIONS PATHWAY
nephrinInteractions<-pathwayList_GenesOneColumn %>% filter(Pathway=="Nephrin interactions")
nephrinInteractions #22 genes total, correct


#PARSE DRUG-TARGET INTERACTION DATA CANCER TARGETOME
#Drug->Target interaction evidence
#Using same version as Gabby's analyses (2017 Targetome evidence file)
allInteractionData<-read.csv(file="Targetome_FullEvidence_070617.txt",header = TRUE,sep = "\t") 

#FILTERING AND CLEAN UP INTERACTION EVIDENCE (same cleaning steps as in TIPS Targetome paper)
#Filter interaction data to target species = human
drugInteractions<-allInteractionData %>% filter(Target_Species=="Homo sapiens")
#Clean up assay types for easier filtering
levels(drugInteractions$Assay_Type)<-list(KD=c("KD", "Kd"), Ki=c("Ki"), IC50=c("EC50"), EC50=c("EC50"), null=c("null"))
#Make sure assay values are read in as numeric!
drugInteractions$Assay_Value<-as.numeric(as.character(drugInteractions$Assay_Value))

#filter to ATC class L drugs= antineoplastics (to match Gabby's analyses for HNSCC light and dark pathways)
#uses accompanying drug information file for drug-target evidence
alldrugData<-read.csv(file = "Targetome_DrugInformation_070617.txt", header= TRUE, sep="\t")
cancerDrugs<-alldrugData%>% filter(grepl("^L", ATC_ClassID)) 
drugInteractions_cleaned<-drugInteractions%>%filter(Drug %in% cancerDrugs$Drug)
#final cleaned targetome evidence file
n_distinct(drugInteractions_cleaned$Drug) #check # of drugs, 137 with evidence, correct (141 total drugs, 137 with interaction evidence)

#GET DRUG-TARGET INTERACTIONS FOR NEPHRIN PATHWAYS
nephrinPathway_Interactions<-drugInteractions_cleaned %>% filter (Target_Name %in% nephrinInteractions$Gene) %>% unique()
  
#Filter interactions to <1000nM, require "=" reported
nephrinPathway_Interactions_1000nM<-nephrinPathway_Interactions %>% filter(Assay_Relation=="=" & Assay_Value<=1000)

#summarize just the best (minimum assay value) for each drug-target interaction
summaryEvidence<-nephrinPathway_Interactions_1000nM %>% 
  group_by(Target_Name, Drug, Assay_Type) %>% 
  summarize(MinAssayValue=min(Assay_Value)) %>%
  select(Target_Name, Drug, Assay_Type, MinAssayValue)
summaryEvidence #interactions we want to visualize

#write table to file
write.table(summaryEvidence, "NephrinInteractions_GeneMembers_withDrugEvidence.txt", sep="\t", row.names = FALSE, col.names = TRUE, quote = FALSE)



```
